---
title: "R-Script Analysis Paper"
output:
  html_document:
    df_print: paged
---

# 1. Fruit trait


## 1.2 Scent

### 1.2.1 GCMS data (variation and PCA)

# In setup-Chunks oben im Notebook
```{r setup, include=FALSE}
#install.packages("here")
library(here)
knitr::opts_knit$set(root.dir = here())
```


```{r}
setwd("C:/Users/frank/OneDrive/Desktop/R_analysis_paper")
```


```{r}
rm(list = ls()) #clean workspace

library(geometry)
library(vegan)
library(readxl)

#Load data

scent_data <- read.csv("data/Scent/raw_table_scent.csv", header=TRUE,sep=";", row.names = 1)
all.species <- unique(scent_data$Species)
print(all.species)
```

Species selection 
```{r}
species_fohanasity <- all.species[2] 
species_guave <- all.species[3]
species_kalafam <- all.species[4]
species_kalafana <- all.species[5]
species_vandrika <- all.species[6]
species_voara <- all.species[7]

```


#### a. Fohanasity

```{r}
# Create dataset and remove metadata
dataset_fohanasity <- subset(scent_data, scent_data$Species == species_fohanasity)
dataset_fohanasity <- dataset_fohanasity[,3:ncol(scent_data)] 
```

Trait variation
```{r}
#Calculate average BC distance in the dataset
fohanasity_mean.bc.dist <- mean(vegdist(dataset_fohanasity, method = "bray")) 
print(fohanasity_mean.bc.dist)
```

Function for finding most inflating sample
```{r}
find.most.inflating.sample <- function(df) {
current.volumes <- c(1:nrow(df)) 
for (j in c(1:nrow(df))) {
  dataset.to.calculate.volume <- df[-j,] 
  hypervolume.df.without.j <- convhulln(as.matrix(dataset.to.calculate.volume),"FA") 
  current.volumes[j] <- hypervolume.df.without.j$vol 
}
    sample.inflating.most <- which(current.volumes == min(current.volumes))
  return(sample.inflating.most)
  }
```

Calculation_PCA
```{r}
pca.dataset_fohanasity <- prcomp(dataset_fohanasity)

#Identify and select PCs that explain at least 90% of the variance
fohanasity_last.pc.to.include <- which(summary(pca.dataset_fohanasity)$importance[3,] > 0.9)[1] 
fohanasity_pcs.to.include <- c(1:fohanasity_last.pc.to.include)
dataset_fohanasity <- pca.dataset_fohanasity$x[,fohanasity_pcs.to.include] 
initial.volume.dataset_fohanasity <- convhulln(as.matrix(dataset_fohanasity),"FA")$vol 
print(dataset_fohanasity)
```
#### b. Guave

```{r}
dataset_guave <- subset(scent_data, scent_data$Species == species_guave) 
dataset_guave <- dataset_guave[,3:ncol(scent_data)] 
```

Trait variation
```{r}
guave_mean.bc.dist <- mean(vegdist(dataset_guave, method = "bray")) 
print(guave_mean.bc.dist)
```

Function for finding most inflating sample
```{r}
find.most.inflating.sample <- function(df) {
current.volumes <- c(1:nrow(df)) 
for (j in c(1:nrow(df))) {
  dataset.to.calculate.volume <- df[-j,] 
  hypervolume.df.without.j <- convhulln(as.matrix(dataset.to.calculate.volume),"FA") 
  current.volumes[j] <- hypervolume.df.without.j$vol 
}
    sample.inflating.most <- which(current.volumes == min(current.volumes))
  return(sample.inflating.most)
  }
```

Calculation_PCA
```{r}
pca.dataset_guave <- prcomp(dataset_guave)

#Identify and select PCs that explain at least 90% of the variance
guave_last.pc.to.include <- which(summary(pca.dataset_guave)$importance[3,] > 0.9)[1] 
guave_pcs.to.include <- c(1:guave_last.pc.to.include)
dataset_guave <- pca.dataset_guave$x[,guave_pcs.to.include]
initial.volume.dataset_guave <- convhulln(as.matrix(dataset_guave),"FA")$vol
print(dataset_guave)
```


#### c. kalafam

```{r}
dataset_kalafam <- subset(scent_data, scent_data$Species == species_kalafam) 
dataset_kalafam <- dataset_kalafam[,3:ncol(scent_data)] 
```

Trait variation
```{r}
kalafam_mean.bc.dist <- mean(vegdist(dataset_kalafam, method = "bray")) 
print(kalafam_mean.bc.dist)
```

Function for finding most inflating sample
```{r}
find.most.inflating.sample <- function(df) {
current.volumes <- c(1:nrow(df)) 
for (j in c(1:nrow(df))) {
  dataset.to.calculate.volume <- df[-j,] 
  hypervolume.df.without.j <- convhulln(as.matrix(dataset.to.calculate.volume),"FA") 
  current.volumes[j] <- hypervolume.df.without.j$vol 
}
    sample.inflating.most <- which(current.volumes == min(current.volumes))
  return(sample.inflating.most)
  }
```

Calculation_PCA
```{r}
pca.dataset_kalafam <- prcomp(dataset_kalafam)

#Identify and select PCs that explain at least 90% of the variance
kalafam_last.pc.to.include <- which(summary(pca.dataset_kalafam)$importance[3,] > 0.9)[1] 
kalafam_pcs.to.include <- c(1:kalafam_last.pc.to.include) 
dataset_kalafam <- pca.dataset_kalafam$x[,kalafam_pcs.to.include] 
initial.volume.dataset_kalafam <- convhulln(as.matrix(dataset_kalafam),"FA")$vol 
print(dataset_kalafam)
```


#### d. kalafana

```{r}
dataset_kalafana <- subset(scent_data, scent_data$Species == species_kalafana) 
dataset_kalafana <- dataset_kalafana[,3:ncol(scent_data)] 
```

Trait-variation
```{r}
kalafana_mean.bc.dist <- mean(vegdist(dataset_kalafana, method = "bray")) 
print(kalafana_mean.bc.dist)
```

Function for finding most inflating sample
```{r}
find.most.inflating.sample <- function(df) {
current.volumes <- c(1:nrow(df)) 
for (j in c(1:nrow(df))) {
  dataset.to.calculate.volume <- df[-j,] 
  hypervolume.df.without.j <- convhulln(as.matrix(dataset.to.calculate.volume),"FA") 
  current.volumes[j] <- hypervolume.df.without.j$vol 
}
    sample.inflating.most <- which(current.volumes == min(current.volumes))
  return(sample.inflating.most)
  }
```

Calculation_PCA
```{r}
pca.dataset_kalafana <- prcomp(dataset_kalafana)

#Identify and select PCs that explain at least 90% of the variance
kalafana_last.pc.to.include <- which(summary(pca.dataset_kalafana)$importance[3,] > 0.9)[1] 
kalafana_pcs.to.include <- c(1:kalafana_last.pc.to.include) 
dataset_kalafana <- pca.dataset_kalafana$x[,kalafana_pcs.to.include] 
initial.volume.dataset_kalafana <- convhulln(as.matrix(dataset_kalafana),"FA")$vol
print(dataset_kalafana)
```


#### e. vandrika

```{r}
dataset_vandrika <- subset(scent_data, scent_data$Species == species_vandrika) 
dataset_vandrika <- dataset_vandrika[,3:ncol(scent_data)] 
```

Trait-variation
```{r}
vandrika_mean.bc.dist <- mean(vegdist(dataset_vandrika, method = "bray")) 
print(vandrika_mean.bc.dist)
```

Function for finding most inflating sample
```{r}
find.most.inflating.sample <- function(df) {
current.volumes <- c(1:nrow(df)) 
for (j in c(1:nrow(df))) {
  dataset.to.calculate.volume <- df[-j,] 
  hypervolume.df.without.j <- convhulln(as.matrix(dataset.to.calculate.volume),"FA") 
  current.volumes[j] <- hypervolume.df.without.j$vol 
}
    sample.inflating.most <- which(current.volumes == min(current.volumes))
  return(sample.inflating.most)
  }
```

Calculation_PCA
```{r}
pca.dataset_vandrika <- prcomp(dataset_vandrika)

#Identify and select PCs that explain at least 90% of the variance
vandrika_last.pc.to.include <- which(summary(pca.dataset_vandrika)$importance[3,] > 0.9)[1] 
vandrika_pcs.to.include <- c(1:vandrika_last.pc.to.include) 
dataset_vandrika <- pca.dataset_vandrika$x[,vandrika_pcs.to.include] 
initial.volume.dataset_vandrika <- convhulln(as.matrix(dataset_vandrika),"FA")$vol 
print(dataset_vandrika)
```


#### f. Voara

```{r}
dataset_voara <- subset(scent_data, scent_data$Species == species_voara) 
dataset_voara <- dataset_voara[,3:ncol(scent_data)] 
```

Trait-variation
```{r}
voara_mean.bc.dist <- mean(vegdist(dataset_voara, method = "bray")) 
print(voara_mean.bc.dist)
```

Function for finding most inflating sample
```{r}
find.most.inflating.sample <- function(df) {
current.volumes <- c(1:nrow(df)) 
for (j in c(1:nrow(df))) {
  dataset.to.calculate.volume <- df[-j,] 
  hypervolume.df.without.j <- convhulln(as.matrix(dataset.to.calculate.volume),"FA") 
  current.volumes[j] <- hypervolume.df.without.j$vol 
}
    sample.inflating.most <- which(current.volumes == min(current.volumes))
  return(sample.inflating.most)
  }
```

Calculation_PCA
```{r}
pca.dataset_voara <- prcomp(dataset_voara)

#Identify and select PCs that explain at least 90% of the variance
voara_last.pc.to.include <- which(summary(pca.dataset_voara)$importance[3,] > 0.9)[1] 
voara_pcs.to.include <- c(1:voara_last.pc.to.include) 
dataset_voara <- pca.dataset_voara$x[,voara_pcs.to.include] 
initial.volume.dataset_voara <- convhulln(as.matrix(dataset_voara),"FA")$vol 
print(dataset_voara)
```


### 1.2.2 Scent-sugar correlation 

```{r}
# Load in data with scent-PCA results and sugar content 

fohanasity <- read.table("data/Scent/Fohanasity_scent_sugar.csv", 
                         sep = ";",
                         header = TRUE)
guave <- read.table("data/Scent/Guave_scent_sugar.csv", 
                         sep = ";",
                         header = TRUE)
kalafam <- read.table("data/Scent/Kalafam_scent_sugar.csv", 
                         sep = ";",
                         header = TRUE)
kalafana <- read.table("data/Scent/Kalafana_scent_sugar.csv", 
                         sep = ";",
                         header = TRUE)
vandrika <- read.table("data/Scent/Vandrika_scent_sugar.csv", 
                         sep = ";",
                         header = TRUE)
voara <- read.table("data/Scent/Voara_scent_sugar.csv", 
                         sep = ";",
                         header = TRUE)


fohanasity$Sample.number <- as.numeric(fohanasity$Sample.number)
guave$Sample.number <- as.numeric(guave$Sample.number)
kalafam$Sample.number <- as.numeric(kalafam$Sample.number)
kalafana$Sample.number <- as.numeric(kalafana$Sample.number)
vandrika$Sample.number <- as.numeric(vandrika$Sample.number)
voara$Sample.number <- as.numeric(voara$Sample.number)


#Calculate correlation between sugar and scent per species
#Fohanasity
scent1_sugar_foha <- cor(fohanasity$PC1, fohanasity$sugar.content)
scent2_sugar_foha <- cor(fohanasity$PC2, fohanasity$sugar.content)
print(scent1_sugar_foha)
print(scent2_sugar_foha)

#guave
scent1_sugar_guave <- cor(guave$PC1, guave$sugar.content)
scent2_sugar_guave <- cor(guave$PC2, guave$sugar.content)
print(scent1_sugar_guave)
print(scent2_sugar_guave)

#kalafam
scent1_sugar_kalafam <- cor(kalafam$PC1, kalafam$sugar.content)
scent2_sugar_kalafam <- cor(kalafam$PC2, kalafam$sugar.content)
print(scent1_sugar_kalafam)
print(scent2_sugar_kalafam)

#kalafana
scent1_sugar_kalafana <- cor(kalafana$PCA1, kalafana$sugar.content)
scent2_sugar_kalafana <- cor(kalafana$PCA2, kalafana$sugar.content)
print(scent1_sugar_kalafana)
print(scent2_sugar_kalafana)

#vandrika
scent1_sugar_vandrika <- cor(vandrika$PC1, vandrika$sugar.content)
scent2_sugar_vandrika <- cor(vandrika$PC2, vandrika$sugar.content)
print(scent1_sugar_vandrika)
print(scent2_sugar_vandrika)

#voara
scent1_sugar_voara <- cor(voara$PC1, voara$sugar.content)
scent2_sugar_voara <- cor(voara$PC2, voara$sugar.content)
print(scent1_sugar_voara)
print(scent2_sugar_voara)

```


## 1.3 Color

PCA calculation 

```{r}

# Guave
spectra_guave <- read.csv("data/Color/spectraldata_guave.csv", sep = ";")
rownames(spectra_guave) <- spectra_guave[, 1] 
spectra_guave <- spectra_guave[, -1]         

# PCA calculation 
pca.guave <- prcomp(spectra_guave, center = T, scale. = T)
summary(pca.guave)

color.guave1 <- pca.guave$x[,1]


# Foha
spectra_foha <- read.csv("data/Color/spectraldata_foha.csv")
rownames(spectra_foha) <- spectra_foha[, 1] 
spectra_foha <- spectra_foha[, -1]  

# PCA calculation 
pca.foha <- prcomp(spectra_foha, center = T, scale. = T)
summary(pca.foha)


color.foha1 <- pca.foha$x[,1]
color.foha2 <- pca.foha$x[,2]

# Kalafana
spectra_kalafana <- read.csv("data/Color/spectraldata_kalafana.csv")
rownames(spectra_kalafana) <- spectra_kalafana[, 1] 
spectra_kalafana <- spectra_kalafana[, -1]  

# PCA calculation 
pca.kalafana <- prcomp(spectra_kalafana[,-1], center = T, scale. = T)
summary(pca.kalafana)


color.kalafana1 <- pca.kalafana$x[,1]
color.kalafana2 <- pca.kalafana$x[,2]


# kalafam
spectra_kalafam <- read.csv("data/Color/spectraldata_kalafam.csv")
rownames(spectra_kalafam) <- spectra_kalafam[, 1] 
spectra_kalafam <- spectra_kalafam[, -1] 

# PCA calculation 
pca.kalafam <- prcomp(spectra_kalafam[,-1], center = T, scale. = T)
summary(pca.kalafam)

color.kalafam1 <- pca.kalafam$x[,1]


# voara
spectra_voara <- read.csv("data/Color/spectraldata_voara.csv")
rownames(spectra_voara) <- spectra_voara[, 1] 
spectra_voara <- spectra_voara[, -1] 

# PCA calculation 
pca.voara <- prcomp(spectra_voara[,-1], center = T, scale. = T)
summary(pca.voara)

color.voara1 <- pca.voara$x[,1]
color.voara2 <- pca.voara$x[,2]

# vandrika
spectra_vandrika <- read.csv("data/Color/spectraldata_vandrika.csv")
rownames(spectra_vandrika) <- spectra_vandrika[, 1] 
spectra_vandrika <- spectra_vandrika[, -1] 

# PCA calculation 
pca.vandrika <- prcomp(spectra_vandrika[,-1], center = T, scale. = T)
summary(pca.vandrika)

color.vandrika1 <- pca.vandrika$x[,1]
color.vandrika2 <- pca.vandrika$x[,2]

#######
#write.csv(color.foha1, file = "PCAs/pca_foha_1.csv")
#write.csv(color.foha2, file = "PCAs/pca_foha_2.csv")
#write.csv(color.guave1, file = "pca_guave_1.csv")
#write.csv(color.kalafam1, file = "PCAs/pca_kalafam_1.csv")
#write.csv(color.kalafana1, file = "PCAs/pca_kalafana_1.csv")
#write.csv(color.kalafana2, file = "PCAs/pca_kalafana_2.csv")
#write.csv(color.vandrika1, file = "PCAs/pca_vandrika_1.csv")
#write.csv(color.vandrika2, file = "PCAs/pca_vandrika_2.csv")
#write.csv(color.voara1, file = "PCAs/pca_voara_1.csv")
#write.csv(color.voara2, file = "PCAs/pca_voara_2.csv")

```



### 1.3.1 Trait variation

```{r, include=FALSE}
# Load in data with Color-PCA and sugar content 

guave_traits <- read.csv("data/Guave_traits.csv")
fohanasity_traits <- read.csv("data/Fohanasity_traits.csv")
kalafana_traits <- read.csv("data/Kalafana_traits.csv")
kalafam_traits <- read.csv("data/Kalafam_traits.csv")
vandrika_traits <- read.csv("data/Vandrika_traits.csv")
voara_traits <- read.csv("data/Voara_traits.csv")


foha_color_pca1 <- read_excel("data/Color/pca_foha_1.xlsx")
foha_color_pca2 <- read_excel("data/Color/pca_foha_2.xlsx")
guave_color_pca1 <- read_excel("data/Color/pca_guave_1.xlsx")
kalafam_color_pca1 <- read_excel("data/Color/pca_kalafam_1.xlsx")
kalafana_color_pca1 <- read_excel("data/Color/pca_kalafana_1.xlsx")
kalafana_color_pca2 <- read_excel("data/Color/pca_kalafana_2.xlsx")
voara_color_pca1 <- read_excel("data/Color/pca_voara_1.xlsx")
voara_color_pca2 <- read_excel("data/Color/pca_voara_2.xlsx")
vandrika_color_pca1 <- read_excel("data/Color/pca_vandrika_1.xlsx")
vandrika_color_pca2 <- read_excel("data/Color/pca_vandrika_2.xlsx")


```

```{r}
cv <- function(x) sd(x)/mean(x)


# Calaculate variation of Color-PCAs

cv_foha_color1 <- cv(foha_color_pca1$PCA1)
cv_foha_color2 <- cv(foha_color_pca2$pca_2)
cv_guave_color1 <- cv(guave_color_pca1$pca_1)
cv_kalafam_color1 <- cv(kalafam_color_pca1$pca_1)
cv_kalafana_color1 <-cv(kalafana_color_pca1$pca_1)
cv_kalafana_color2 <-cv(kalafana_color_pca2$pca_2)
cv_vandrika_color1 <-cv(vandrika_color_pca1$pca_1)
cv_vandrika_color2 <-cv(vandrika_color_pca2$pca_2)
cv_voara_color1 <-cv(voara_color_pca1$pca_1)
cv_voara_color2 <-cv(voara_color_pca2$pca_2)
```


### 1.3.2 Color-sugar correlation 

Correlation 

```{r}

#Calculate correlation of Color-PCA and sugar content per species
color1_sugar_guave <- cor(guave_traits$color_pca1, guave_traits$sugar.content)

color1_sugar_foha <- cor(fohanasity_traits$color_pca1, fohanasity_traits$sugar.content)
color2_sugar_foha <- cor(fohanasity_traits$color_pca2, fohanasity_traits$sugar.content)

color1_sugar_kalafam <- cor(kalafam_traits$color_pca1, kalafam_traits$sugar.content)

color1_sugar_kalafana <- cor(kalafana_traits$color_pca1, kalafana_traits$sugar.content)
color2_sugar_kalafana <- cor(kalafana_traits$color_pca2, kalafana_traits$sugar.content)

color1_sugar_voara <- cor(voara_traits$color_pca1, voara_traits$sugar.content)
color2_sugar_voara <- cor(voara_traits$color_pca2, voara_traits$sugar.content)

color1_sugar_vandrika <- cor(vandrika_traits$color_pca1, vandrika_traits$sugar.content)
color2_sugar_vandrika <- cor(vandrika_traits$color_pca2, vandrika_traits$sugar.content)
```


## 1.4 Hardness

### 1.4.1 Trait variation 

```{r}
cv <- function(x) sd(x)/mean(x)

# Calculate variation in hardness per species 

cv_foha_hardness <- cv(fohanasity_traits$value.puntc..res)
cv_guave_hardness <- cv(guave_traits$value.puntc..res)
cv_kalafam_hardness <- cv(kalafam_traits$value.puntc..res)
cv_kalafana_hardness <-cv(kalafana_traits$value.puntc..res)
cv_vandrika_hardness <-cv(vandrika_traits$value.puntc..res)
cv_voara_hardness <-cv(voara_traits$value.puntc..res)
```


### 1.4.2 Hardness-sugar variation 

```{r}
# Calculate Correlation between hardness and sugar content per species 

cor(fohanasity_traits$value.puntc..res, fohanasity_traits$sugar.content)
cor(guave_traits$value.puntc..res, guave_traits$sugar.content)
cor(kalafam_traits$value.puntc..res, kalafam_traits$sugar.content)
cor(kalafana_traits$value.puntc..res, kalafana_traits$sugar.content)
cor(vandrika_traits$value.puntc..res, vandrika_traits$sugar.content)
cor(voara_traits$value.puntc..res, voara_traits$sugar.content)

```

# 2. Analysis 


Installing and loading packages
```{r results='hide', warning=FALSE, message=FALSE}
#install.packages("lme4")
library("lme4")
```

Test for overdispersion (by Roger Mundry)
```{r results='hide'}
overdisp.test<-function(x){
  pr=residuals(x, type ="pearson")
  sum.dp=sum(pr^2)
  if(class(x)[[1]]=="mer"){
    xdf=length(residuals(x))-length(fixef(x))
  }else if(class(x)[[1]]=="glmerMod"){
    xdf=length(residuals(x))-length(fixef(x))
  }else{
    xdf=length(residuals(x))-length(x$coefficients)
  }
  return(data.frame(chisq=sum.dp, df=xdf, P=1-pchisq(sum.dp, xdf), dispersion.parameter=sum.dp/xdf))
}
```

Load and modify data
```{r results='hide'}
dataset <- read.table("data/Data_mixed_model.csv", 
                      sep = ",", 
                      header = TRUE)

dataset$look.yes <- ifelse(dataset$Behaviour.code == 'L', 1, 0)
dataset$sniff.yes <- ifelse(dataset$Behaviour.code == 'S', 1, 0)
dataset$touch.yes <- ifelse(dataset$Behaviour.code == 'T', 1, 0)

dataset <- subset(dataset, dataset$Behaviour.code != "R")

dataset$variation.colour1 <- scale(dataset$variation.colour1)
dataset$variation.scent <- scale(dataset$variation.scent)
dataset$variation.hardness <- scale(dataset$variation.hardness)

dataset$variation.colour1 <- as.numeric(dataset$variation.colour1)
dataset$variation.scent <- as.numeric(dataset$variation.scent)
dataset$variation.hardness <- as.numeric(dataset$variation.hardness)
```

GLMM for looking behaviour
```{r results='hide', warning=FALSE, message=FALSE}
table(dataset$look.yes)

#variation colour
look.var.all <- glmer(look.yes~variation.colour1+
                         (1|Tree.Local.name)+
                         (1|Individuum.ID/Group.ID), 
                       data = dataset, 
                       family = binomial)

#correlation colour and sugar 
look.cor.all <- glmer(look.yes~abs(colour1.sugar)+ 
                         (1|Tree.Local.name)+
                         (1|Individuum.ID/Group.ID), 
                       data = dataset, 
                       family = binomial)
```

Output for looking behaviour
```{r}
#Intraspecific variation
overdisp.test(look.var.all)
summary(look.var.all)

#Colour-sugar correlation
overdisp.test(look.cor.all)
summary(look.cor.all)
```

GLMM for sniffing behaviour
```{r results='hide', warning=FALSE, message=FALSE}
table(dataset$sniff.yes)

#variation scent 
sniff.var.all <- glmer(sniff.yes~variation.scent+
                     (1|Tree.Local.name)+
                     (1|Individuum.ID/Group.ID), 
                   data = dataset, 
                   family = binomial)

#correlation scent and sugar
sniff.cor.all <- glmer(sniff.yes~abs(scent1.sugar)+abs(scent2.sugar)+
                     (1|Tree.Local.name)+
                     (1|Individuum.ID/Group.ID), 
                   data = dataset, 
                   family = binomial)
```

Output for sniffing behaviour
```{r}
#Intraspecific variation
overdisp.test(sniff.var.all)
summary(sniff.var.all)

#Scent-sugar correlation
overdisp.test(sniff.cor.all)
summary(sniff.cor.all)
```

GLMM for touching behaviour
```{r results='hide', warning=FALSE, message=FALSE}
table(dataset$touch.yes)

#variation hardness
touch.var.all <- glmer(touch.yes~variation.hardness+
                         (1|Tree.Local.name)+
                         (1|Individuum.ID/Group.ID), 
                       data = dataset, 
                       family = binomial)

#correlation hardness and sugar
touch.cor.all <- glmer(touch.yes~abs(hardness.sugar)+
                         (1|Tree.Local.name)+
                         (1|Individuum.ID/Group.ID), 
                       data = dataset, 
                       family = binomial)
```

Output for touching behaviour
```{r}
#Intraspecific variation
overdisp.test(touch.var.all)
summary(touch.var.all)

#Hardness-sugar correlation
overdisp.test(touch.cor.all)
summary(touch.cor.all)
```

Visualisation: 

Colouring for groups and individuals 
```{r results='hide'}
group_colors <- c("red", "blue", "green")
# Farbpalette für die Individuen festlegen
individual_colors <- c("#FF0000","#8B1C62","red4",  
                       "lightblue", "#00F5FF","royalblue2", "darkblue",
                       "lightgreen", "darkgreen")

par(mar = c(5, 5, 4, 4) + 0.1, oma = c(0, 1, 2, 0))
par(font.axis = 2)
```
`
Visualisation colour
```{r}
#variation colour
plot(dataset$variation.colour1, jitter(dataset$look.yes), yaxt = "n",
     ylab = "", xlab = "Intraspecific variation of colour", cex.lab = 1,
     col = individual_colors[as.factor(dataset$Individuum.ID)], pch = 19, cex = 0.9)
title(main = "Looking vs. variation in fruit colour", cex.main = 1.2)
axis(2, at = c(0, 1), labels = c("No look", "Look"), las = 2, cex.axis = 0.9)
legend("right", legend = unique(dataset$Group.ID), fill = group_colors, title = "Group",
       x = "right", y = 1.05, cex = 0.6)

xvals.var.col <- seq(from = min(dataset$variation.colour1), to = max(dataset$variation.colour1), length.out = 100)
LP.var.col <- fixef(look.var.all)[1] + xvals.var.col*fixef(look.var.all)[2]
yvals.var.col <- exp(LP.var.col) / (1+exp(LP.var.col))
lines(xvals.var.col, yvals.var.col, col = "black", lwd = 2.5)


#correlation colour sugar 
plot(abs(dataset$colour1.sugar), jitter(dataset$look.yes), yaxt = "n",
     ylab = "", xlab = "Correlation colour [absolute] with sugar",
     col = individual_colors[as.factor(dataset$Individuum.ID)], pch = 19, cex = 0.9) 
title(main = "Looking vs. colour-sugar correlation", cex.main = 1.2)
axis(2, at = c(0, 1), labels = c("No look", "Look"), las = 2, cex.axis = 0.9) 
legend("right", legend = unique(dataset$Group.ID), fill = group_colors, title = "Group",
       x = "right", y = 1.05, cex = 0.6)

xvals.cor.col <- seq(from = min(abs(dataset$colour1.sugar)), to = max(abs(dataset$colour1.sugar)), length.out = 100)
LP.cor.col <- fixef(look.cor.all)[1] + xvals.cor.col*fixef(look.cor.all)[2]
yvals.cor.col <- exp(LP.cor.col) / (1+exp(LP.cor.col))
lines(xvals.cor.col, yvals.cor.col, col = "black", lwd = 2.5)
```

Visualisation scent 
```{r}

#variation scent 
plot(dataset$variation.scent, jitter(dataset$sniff.yes), yaxt = "n",
     ylab = "", xlab = "Intraspecific variation of scent", cex.lab = 1,
     col = individual_colors[as.factor(dataset$Individuum.ID)], pch = 19, cex = 0.9)
title(main = "Sniffing vs. variation in fruit scent", cex.main = 1.2)
axis(2, at = c(0, 1), labels = c("No sniff", "Sniff"), las = 2, cex.axis = 0.9)
legend("right", legend = unique(dataset$Group.ID), fill = group_colors, title = "Group",
       x = "right", y = 1.05, cex = 0.6)

xvals.var.scent <- seq(from = min(dataset$variation.scent), to = max(dataset$variation.scent), length.out = 100)
LP.var.scent <- fixef(sniff.var.all)[1] + xvals.var.scent*fixef(sniff.var.all)[2]
yvals.var.scent <- exp(LP.var.scent) / (1+exp(LP.var.scent))
lines(xvals.var.scent, yvals.var.scent, col = "black", lwd = 2.5)


#correlation scent1 sugar 
plot(abs(abs(dataset$scent1.sugar)), jitter(dataset$sniff.yes), yaxt = "n",
     ylab = "", xlab = "Correlation scent [absolute] with sugar",
     col = individual_colors[as.factor(dataset$Individuum.ID)], pch = 19, cex = 0.9) 
title(main = "Sniffing vs. scent-sugar correlation (PC1)", cex.main = 1.2)
axis(2, at = c(0, 1), labels = c("No sniff", "Sniff"), las = 2, cex.axis = 0.9) 
legend("right", legend = unique(dataset$Group.ID), fill = group_colors, title = "Group",
       x = "right", y = 1.05, cex = 0.6)

xvals.cor.scent1 <- seq(from = min(abs(dataset$scent1.sugar)), to = max(abs(dataset$scent1.sugar)), length.out = 100)
LP.cor.scent1 <- fixef(sniff.cor.all)[1] + xvals.cor.scent1*fixef(sniff.cor.all)[2]
yvals.cor.scent1 <- exp(LP.cor.scent1) / (1+exp(LP.cor.scent1))
lines(xvals.cor.scent1, yvals.cor.scent1, col = "black", lwd = 2.5)

#correlation scent2 sugar 
plot(abs(dataset$scent2.sugar), jitter(dataset$sniff.yes), yaxt = "n",
     ylab = "", xlab = "Correlation scent [absolute] with sugar",
     col = individual_colors[as.factor(dataset$Individuum.ID)], pch = 19, cex = 0.9) 
title(main = "Sniffing vs. scent-sugar correlation (PC2)", cex.main = 1.2)
axis(2, at = c(0, 1), labels = c("No sniff", "Sniff"), las = 2, cex.axis = 0.9) 
legend("right", legend = unique(dataset$Group.ID), fill = group_colors, title = "Group",
       x = "right", y = 1.05, cex = 0.6)

xvals.cor.scent2 <- seq(from = min(abs(dataset$scent2.sugar)), to = max(abs(dataset$scent2.sugar)), length.out = 100)
LP.cor.scent2 <- fixef(sniff.cor.all)[1] + xvals.cor.scent2*fixef(sniff.cor.all)[3]
yvals.cor.scent2 <- exp(LP.cor.scent2) / (1+exp(LP.cor.scent2))
lines(xvals.cor.scent2, yvals.cor.scent2, col = "black", lwd = 2.5)
```

Visualisation hardness
```{r}

#variation hardness
plot(dataset$variation.hardness, jitter(dataset$touch.yes), yaxt = "n",
     ylab = "", xlab = "Intraspecific variation of hardness", cex.lab = 1,
     col = individual_colors[as.factor(dataset$Individuum.ID)], pch = 19, cex = 0.9)
title(main = "Touching vs. variation in fruit hardness", cex.main = 1.2)
axis(2, at = c(0, 1), labels = c("No touch", "Touch"), las = 2, cex.axis = 0.9)
legend("right", legend = unique(dataset$Group.ID), fill = group_colors, title = "Group",
       x = "right", y = 1.05, cex = 0.6)

xvals.var.hard <- seq(from = min(dataset$variation.hardness), to = max(dataset$variation.hardness), length.out = 100)
LP.var.hard <- fixef(touch.var.all)[1] + xvals.var.hard*fixef(touch.var.all)[2]
yvals.var.hard <- exp(LP.var.hard) / (1+exp(LP.var.hard))
lines(xvals.var.hard, yvals.var.hard, col = "black", lwd = 2.5)

#correlation hardness sugar
plot(abs(abs(dataset$hardness.sugar)), jitter(dataset$touch.yes), yaxt = "n",
     ylab = "", xlab = "Correlation hardness [absolute] with sugar",
     col = individual_colors[as.factor(dataset$Individuum.ID)], pch = 19, cex = 0.9) 
title(main = "Touching vs. hardness-sugar correlation", cex.main = 1.2)
axis(2, at = c(0, 1), labels = c("No touch", "Touch"), las = 2, cex.axis = 0.9) 
legend("right", legend = unique(dataset$Group.ID), fill = group_colors, title = "Group",
       x = "right", y = 1.05, cex = 0.6)

xvals.cor.hard <- seq(from = min(abs(dataset$hardness.sugar)), to = max(abs(dataset$hardness.sugar)), length.out = 100)
LP.cor.hard <- fixef(touch.cor.all)[1] + xvals.cor.hard*fixef(touch.cor.all)[2]
yvals.cor.hard <- exp(LP.cor.hard) / (1+exp(LP.cor.hard))
lines(xvals.cor.hard, yvals.cor.hard, col = "black", lwd = 2.5)
```

